---
title: "CIND820 Capstone"
author: "Marlene Martins"
date: "March 29, 2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


install.packages('FNN') 
install.packages('RCurl') 
install.packages('MASS') 
install.packages('leaps') 
library(MASS) # stepwise regression 
library(leaps) # all subsets regression

Dataset: http://vincentarelbundock.github.io/ Rdatasets/csv/Ecdat/Computers.csv



```{r step1}
library(MASS) # stepwise regression 
library(leaps) # all subsets regression
library(RCurl)


#c_prices <- read.csv(file="http://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/Computers.csv")
covid_dat <- read.csv(file="Q:/BI Team/Credit Card Reporting/Projects/ADH0820/Data/COVID_NBH_Summaryo.csv")

head(covid_dat)
summary(covid_dat)
str(covid_dat)

#solution: doesn't work on my version
#u <- getURL("http://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/Computers.csv") 
#c_prices <- read.csv(text = u)



```


1. Create a multiple linear regression model to predict infection rate.
(a) Split the dataset to 70% of training and 30% of test sets. 




```{r step2}

#The floor function rounds a numeric input down to the next lower integer
rn_train <- sample(nrow(covid_dat), floor(nrow(covid_dat)*0.7)) 
train <- covid_dat[rn_train,] 
test <- covid_dat[-rn_train,]

str(covid_dat)

```

(b) Create the multiple regression model on the train set using ram, screen, speed, hd and ads as independent variables.

We will train our model on the training set and do the prediction on the test set

```{r step3}

model_mlr<- lm(INFECTION_RATE~P_AGE_40_to_64+P_VM_Yes+P_EDU_HS_Lower+P_HH_5_Persons+P_OCC_Ess_Yes, data=train)
model_mlr


```

(c) Apply predict() function on the test set. 

```{r step4}

prediction <- predict(model_mlr, interval="prediction", newdata =test)

head(prediction)
summary(prediction)
```
(d) Calculate the errors (prediction price – test price) in predictions and show the histogram of errors. 

Get difference between prediction and test 
```{r step5}

#this option for prediction "fit" model? 
#get the errors
#this "fit" is a column, the fitted value/model.  lower/upper are intervals upper/lower limit, we want to see model prediction.  we want to compare the fitted value to the test
summary(prediction)
errors <- prediction[,"fit"] - test$INFECTION_RATE 
hist(errors)

# The Error Histogram view can help you to understand how error is distributed
#histogram is left scewed, 


errors2 <- prediction - test$price


```
(e) Compute the root mean squared error.

need to get the error and then square it and get the summation and divide by number of observations in test set
comparing the actual price to the predicted price

```{r step6}

rmse <- sqrt(sum((prediction[,"fit"]-test$INFECTION_RATE)^2)/nrow(test)) 
rmse

#both pred(25) and rmse values are better for multiple linear regression
```


##2. Apply forward selection and backward elimination algorithms to the dataset.

Forward selection:

we end up using all variables

```{r step7}

#Forward SElection

# for all variables


full <- lm(INFECTION_RATE~P_AGE_40_to_64+P_VM_Yes+P_EDU_HS_Lower+P_HH_5_Persons+P_OCC_Ess_Yes,data=covid_dat) 

# this creates for only one independent value at a time
null <- lm(INFECTION_RATE~1,data=covid_dat) 

#stepAIC 
stepF <- stepAIC(null, scope=list(lower=null, upper=full), direction= "forward", trace=TRUE) 

summary(stepF)

```

2. Apply backward elimination algorithms to the dataset.


```{r step8}
#backward elimination

full <- lm(INFECTION_RATE~P_AGE_40_to_64+P_VM_Yes+P_EDU_HS_Lower+P_HH_5_Persons+P_OCC_Ess_Yes,data=covid_dat) 

stepB <- stepAIC(full, direction= "backward", trace=TRUE) 

summary(stepB) 




```

3. Apply regsubsets() to see the best combination of the attributes.  

In the output, * denotes the variables included to the subset. For example, the best combination of 4 attributes is: ram, speed, screen and trend, and the best combination of 5 attributes is: ram, hd, speed, screen and trend

This algorithm uses AIC at the back.  putting all the combinatons and seeing which one has lower AIC to be better.  
part of a package.  
```{r step9}

# specify the model that is being used
subsets<-regsubsets(INFECTION_RATE~P_AGE_40_to_64+P_VM_Yes+P_EDU_HS_Lower+P_HH_5_Persons+P_OCC_Ess_Yes, data=covid_dat, nbest=1) 

sub.sum <- summary(subsets) 

as.data.frame(sub.sum$outmat) 


```

Random Forest 





```{r step10}

library(rsample)
library(randomForest)
#library(ranger)
#library(caret)
#library(h2o)

#set random seed to make results reproducible

set.seed(17)

#calculate the size of each of tre data sets;
data_set_size <- floor(nrow(covid_dat)/2)

#generate a random sample of "data_set_size" indexes
indexes <- sample(1:nrow(covid_dat), size = data_set_size)

#assign the data to the correct sets
training = covid_dat[indexes,]
validation1 = covid_dat[-indexes,]


# Create training (70%) and test (30%) sets for the AmesHousing::make_ames() data.
# Use set.seed for reproducibility

set.seed(123)
data_split <- initial_split(covid_dat, prop = .7)
data_train <- training(data_split)
data_test  <- testing(data_split)


```


```{r step11}

#import the package
library(randomForest)

class(covid_dat$INFECTION_RATE2)
#perform training'

INFECTION_RATE3<-factor(data_train$INFECTION_RATE2)
class(INFECTION_RATE3)

rf_infect_rate <- randomForest(INFECTION_RATE ~P_AGE_40_to_64+P_VM_Yes+P_EDU_HS_Lower+P_HH_5_Persons+P_OCC_Ess_Yes, data=train, ntree=27)

print(rf_infect_rate)


```

```{r step12}
#Run on testing
library(MASS)
library(ModelMetrics)

res = lapply(c(111,222),function(i){
  set.seed(i)

fit = randomForest(INFECTION_RATE ~P_AGE_40_to_64+P_VM_Yes+P_EDU_HS_Lower+P_HH_5_Persons+P_OCC_Ess_Yes, data=train, ntree=27)
  
pred_values = predict(fit,test)
actual_values = test$INFECTION_RATE

#rmse(pred_values,actual_values)

data.frame(seed=i,
      metrics_rmse = rmse(pred_values,actual_values),
      cal_rmse=mean((pred_values-actual_values)^2)^0.5
     )
})
res=do.call(rbind,res)
head(res)


```

```{r lab13}

# Output to be present as PNG file 
png(file = "Q:/BI Team/Credit Card Reporting/Projects/ADH0820/Charts/randomForestRegression.png")
  
# Plot the error vs the number of trees graph
plot(rf_infect_rate)


# Saving the file
dev.off()

varImpPlot(rf_infect_rate)



``` 